Elimination Of Jump Objects

### 2022-09-20 paper/skeleton.tex
upd paper

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ffcode}
\usepackage{amsmath}

\newtheorem{definition}{Definition}[section]

\graphicspath{ {./images/} }

@@

\keywords{Object-Oriented Programming, software engineering, algorithms, computer science}

\section{Introduction}
The problem is known both computer science and soft-
ware engineering communities for a few decades and formulated by several scientists.

Many of them have offered solutions to this problem. (Look at Related Work)

The presence of jump statements in the source code makes it impossible to use the functional method of analysis.

There are an "object-flow" languages where we can control flow through an objects. An example of such language is EO (Elementary Objects language). This is a strictly object-oriented programming language in which any action is described by creating a new object, or decorating an existing one.

We are trying to answer the following research question: "Is it possible for any program, written on "object-flow" language like EO and containing GOTO objects to find a semantically equivalent program, but without using them?"

\section{Related work}
To solve the problem of eliminating Jump statements in other programming languages, the following works were published:
Morris suggested to use regular expressions for replacing GOTO with IF-THEN-ELSE constructs;
Ramshow proposed an algorithm for Pascal programs;
Ceccato suggested how to eliminate GOTO during migration of legacy code to Java.

To our knowledge, there is still no method which enables automatic elimination of jump objects in programs written in such an object-flow languages.

@@

Some objects in EO programs may need to be platform specific and can’t be composed from other existing objects—they are called \emph{atoms}. 
So all jump statements are implemented as a
single \emph{atom}-object GOTO. The implementation of goto-object in EO implies two types of jumps - goto forward and goto backward. So we should consider transformations for both cases.

Since we consider a solution to the problem in strictly object-oriented programming languages, which means that the whole program is a set of objects where some objects can contain others.
It should be noted that each object in the program has its own level of \emph{nesting} relative to other objects.

\begin{definition} The \emph{nesting} of the current object means the number of objects that contain the current object directly, or that contain objects that contain the current object.
\end{definition}

The solution we propose contains two types of transformations - forward jump and backward jump.
The algorithm itself consists in applying these transformations to replace both types of jumps.

\vspace{5mm}

\emph{Designations:}

\begin{itemize}
\item $\sigma$ - Denotes either an object, a chain of nested objects or the absence of an object.

\item $\delta$ - Denotes a not-empty object or a chain of nested objects that return a not-NULL value when accessed.

\item $\mu$ - Denotes an object or a chain of nested objects that return a boolean value (TRUE/FALSE) when executed.
\end{itemize}

We will also look at our transformations in the form of $\varphi$-calculus, on which EO is based. The proposed $\varphi$-calculus represents an object model through data and objects, while operations with them are possible through abstraction, application, and decoration.
The calculus introduces a formal apparatus for manipulations with objects.

$\varphi$-calculus uses an alternative “arrow notation” to denote an object in a more compact way, where free attributes of an abstract object stay in the parentheses on the left side of the mapping symbol $\mapsto$ and pairs, which represent bound attributes, stay on the right side, in double-square brackets.

\subsection{Forward jump}
Goto forward in EO is implemented as:

\begin{ffcode}
+import org.eolang.gray.goto
+import org.eolang.io.stdout

goto
  [g]
    |\sgm{1}| > @
      g.forward |\dlt{1}|
      stdout "Will never be printed"
\end{ffcode}

It consists of the goto atom-object, the abstract object \emph{g}, and the g.forward attribute. Several abstract objects such as \emph{g} can be declared inside of goto-object, and the attribute \emph{.forward} can be called to each of them.

\subsubsection{Simple Forward}
Consider the simplest example of using goto-forward:

\begin{ffcode}
goto
  [g]
    g.forward |\dlt{1}| > @
\end{ffcode}

In terms of $\varphi$-calculus, this code will appear as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \ff{g}.\ff{forward} (\delta_1) \rrbracket ) \\
\end{split}
\end{equation}

Note that in this example, g.forward is simply an attribute of the abstract object \emph{g} and is executed as soon as the queue reaches it.
Our transformation only considers the jumps that are made when a condition for it is fulfilled. In cases where there is unconditional jump, first of all we transform it into conditional jump by adding an additional if-statement - if (TRUE).
This will not violate the program logic, because without the added if-statement, the jump would always be executed, and statements coming after this jump with a nesting not less that the nesting of the abstract object \emph{g} would never be executed:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.( TRUE, \ff{g}.\ff{forward}(\delta_1), TRUE ) \rrbracket) 
\end{split}
\end{equation}

Same example on EO:

\begin{ffcode}
goto
  [g]
    if. > @
      TRUE
      g.forward |\dlt{1}|
      TRUE
\end{ffcode}

\subsubsection{Goto forward wrapped in an object}

After the conversion described in section 3.1.1, the simplest call of forward jump will be presented as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

The transformation under consideration involves the following steps:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1.\ff{not}, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1.\ff{not}, \delta_2, \delta_1) \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{if.}(\mu_1.not, \delta_2, \delta_1) \\
\end{split}
\end{equation}

\begin{theorem}
Expression 3 is semantically equivalent to expression 6:
\end{theorem}

\begin{equation} \nonumber
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

\hline

\begin{equation} \nonumber
\begin{split}
& \ff{if.}(\mu_1.not, \delta_2, \delta_1) \\
\end{split}
\end{equation}

To prove that the transformation does not violate program logic, let's consider it with the following abstract case of calling goto forward.

Consider the following code:

\begin{ffcode}
goto
  [g]
    |\sgm{1}| > @ |$\label{ln:ret0}$|
      |\sgm{2}|
      if.
        |\m{1}|
        g.forward |\dlt{1}|
        |\dlt{2}|
      |\sgm{3}| |$\label{ln:ret1}$|
\end{ffcode}

Let's denote lines \ref{ln:ret0}-\ref{ln:ret1} as X.

It is not difficult to notice that statements which goto-forward "jumps through" will be executed if the condition in the if-statement is not equal to TRUE.
Otherwise statements following the goto-object itself are executed.
So we can swap statements and invert the condition into if-statement to get rid of the \emph{g.forward} attribute without violating program logic.

Also, after replacing the jump, we must create an additional \emph{flag} object, which will be changed if the jump condition is met, i.e. when in the new if-statement, the 'ELSE' branch will be executed - value of flag will change. This is required so that all objects following the if-statement that are attributes of the abstract object g are not executed.
So, by wrapping each object following the jump condition and whose nesting is at least that of the g object in an if-statement with a flag value check condition, we make sure that these objects will not be executed when the jump condition is met.

By swapping the statements that follow the if-statement body and after the jump itself, as described above, we get the following replacement for this example: the \m{1} object must be inverted with (\m{1}).not; in 'THEN' brunch of if-statement places \dlt{2}; in the 'ELSE' brunch creates new object seq, consisting of two elements - \dlt{1} and flag's value change, because the jump condition is satisfied.

Thus, after transformation, the program converts to:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}||$\label{ln:ret2}$|
  |\sgm{2}|
  if.
    |\m{1}|.not
    |\dlt{2}|
    seq
      |\dlt{1}|
      flag.write 1
  if.
    (eq. (flag 1)).not
    |\sgm{3}|
    TRUE|$\label{ln:ret3}$|
\end{ffcode}

Let's denote lines \ref{ln:ret2}-\ref{ln:ret3} as X'.

Now let's look at an example program with nested objects containing goto forward:
\begin{ffcode}
goto
  [g]
    |\sgm{0}| > @
      X
      |\sgm{4}|
\end{ffcode}

Applying the transformation, we get:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}|
  X'
  if.
    (eq. (flag 1)).not
    |\sgm{4}|
    TRUE
\end{ffcode}

\subsubsection{Goto forward inside the while-loop}
In order to realize the functionality of cycles in EO, generators are provided. A generator is a routine that can be used to control the iteration behaviour of a loop.

The language currently provides a while generator, on which other generators are implemented.

The simplest way to call goto forward inside a while-loop is represented as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
& \quad \quad \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

The transformation under consideration involves the following steps:

%%\begin{equation}
%%\begin{split}
%%& \ff{memory}(flag), \\
%%& flag.\ff{write}(0), \\
%%& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} %%\mapsto \varnothing, \\
%%& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
%%& \quad \ff{if}.(\mu_1, %%\ff{g}.\ff{forward}(\delta_1), \delta_2) \\
%%& \quad )\rrbracket) \\
%%\end{split}
%%\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{and}.( \\
& \quad \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \quad \ff{if}.(\mu_1,  \ff{g}.\ff{forward}(\delta_1), \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{and}.( \\
& \quad \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \quad \ff{if}.(\mu_1.\ff{not}, \delta_2, \\
& \quad \quad \quad \ff{seq}( \delta_1, flag.\ff{write}(0) )\\
& \quad \quad ) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{and}.( \\
& \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \ff{if}.(\mu_1.\ff{not}, \delta_2, \\
& \quad \quad \ff{seq}( \delta_1, flag.\ff{write}(0) )\\
& \quad ) \\
& ) \\
\end{split}
\end{equation}

\begin{theorem}
Expression 7 is semantically equivalent to expression 10:
\end{theorem}

\begin{equation} \nonumber
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
& \quad \quad \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\hline

\begin{equation} \nonumber
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{and}.( \\
& \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \ff{if}.(\mu_1.\ff{not}, \delta_2, \\
& \quad \quad \ff{seq}( \delta_1, flag.\ff{write}(0) \\
& \quad )\\
& ) \\
\end{split}
\end{equation}

To prove that the transformation does not violate program logic, let's consider it with the following abstract case of calling goto forward.

Consider the following code on EO:

\begin{ffcode}
goto
  [g]
    |\sgm{0}| > @ |$\label{ln:ret4}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.forward |\dlt{1}|
            |\dlt{2}|
          |\sgm{3}||$\label{ln:ret5}$|
\end{ffcode}

>>>>

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false,backend=biber]{biblatex}
\usepackage{debate}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
\renewcommand\ttdefault{cmtt}
\usepackage{csquotes}
\usepackage{ffcode}
\usepackage{CJKutf8}
\usepackage{paralist}
\usepackage{cancel} % to enable \cancel command
\usepackage{anyfontsize} % To get rid of font not found warnings
\usepackage{tabularx} % for special tables
\usepackage{amsmath}
\usepackage{to-be-determined}
\usepackage{href-ul}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{tcolorbox} % for algorithm
\tcbuselibrary{skins}
\usepackage{fdsymbol} % for \mathbb
\usepackage{algpseudocode} % for algorithms
\usepackage{multicol} % for two cols in BNF
\usepackage{pgffor} % to enable \foreach
\usepackage{stmaryrd}
\usepackage{mathtools}
\newtheoremstyle{theorems}
{} % Space above
{} % Space below
{} % Theorem body font % (default is "\upshape")
{} % Indent amount
{\scshape} % Theorem head font % (default is \mdseries)
{.} % Punctuation after theorem head % default: no punctuation
{ } % Space after theorem head
{} % Theorem head spec
\theoremstyle{theorems}
\newtheorem{eodefinition}{Definition}
\newtheorem{eotheorem}{Theorem}

\addbibresource{refs.bib}

\newcommand\br{\\[-4pt]}
\newcounter{rule}
\renewcommand\therule{R\arabic{rule}}
\newcommand\rr{\smalltriangleright{}}
\newcommand\rrule[1]{{\scshape\sffamily\ref{rule:#1}}}
\newcommand{\jrule}[1]{%
    \refstepcounter{rule}\label{rule:#1}%
    \text{\textbf{\rrule{#1}}}}

@@

    \section{Introduction}
    The problem of eliminating Jump-like statements is known for both computer science and software engineering communities for a few decades and formulated by several scientists, some of which have suggested solutions for this problem.

    The presence of jump statements in the source code makes it impossible to use the functional method of analysis.
    Other program analysis methods, including those based on symbolic execution techniques, will also benefit from the removal of jump statements, which are well-known obstacles to the precision of analysis, as explained by \citet{baldoni2018survey}.

    There are an "object-flow" languages where the execution flow is controlled by object composition.
    An example of such a language is EO, which is based on $\varphi$-calculus introduced by \citet{bugayenko2021eolang} and formalized by \citet{kudasov2021}.
    This is a strictly object-oriented programming language in which any action is described by creating a new object, or decorating an existing one.

    We are trying to answer the following research question: "Is it possible for any program, written in 'object-flow' language like EO and containing GOTO objects, to find a semantically equivalent program, but without GOTO objects?"

    \section{Related work}
    The problem was firstly formulated by \citet{dijkstra1968letters}: he affirmed that the use of GOTO has disastrous effect and proposed to remove it from all programming languages.
    Even though a few years later \citet{knuth1974structured} showed a manner for structuring programs with the GOTO statement and argued that it could be a powerful tool if it is well used, researchers got interested in finding a way for restructuring programs and eliminating GOTO statements \citep{arsac1977construction}.

    To solve the problem of eliminating Jump statements in other programming languages, the following works were published:
    \Citet{williams1985restructuring} created a tool for restructuring Pascal code;
    \Citet{jonsson1989next} was one of the first who made a summary of so called GOTO-patches, which are design techniques for eliminating GOTO statements;
    \Citet{erosa1994taming} suggested an algorithm for automatic elimination of GOTO statements in~C;
    \Citet{zegour1994new} explained how GOTO elimination may work in any language, as long as it has jump statements and control structures;
    \Citet{morris1997goto} suggested to use regular expressions for replacing GOTO with IF-THEN-ELSE constructs;
    \Citet{ramshaw1988eliminating} proposed an algorithm for Pascal programs;
    \Citet{ganapathi2008shim} introduced optimization method for SHIM IR;
    \Citet{ceccato2008goto} suggested how to eliminate GOTO during migration of legacy code to Java.

    However, to our knowledge, there is still no method available, which enables automatic elimination of jump objects in programs written in such an object-flow languages.

    \section{Definitions and designations}
    \begin{eodefinition}
        \emph{Dataization} in $\varphi$-calculus is a process of retrieving data from an object.
        The dataization of an object at the highest level of composition leads to the execution of a program.
    \end{eodefinition}
    \begin{eodefinition} \emph{Nesting} of the current object determines the number of objects that encapsulate current object via abstraction, application, or decoration directly or through other objects.

    For example, here object $x$ has 3-rd level of nesting:
    \begin{ffcode}
        o1
        o2
        ...
        o3
        o4
        x
    \end{ffcode}
    \end{eodefinition}
    \begin{eodefinition}
        \emph{Forwarding goto} denotes an object \ff{goto}, which encapsulate object \ff{forward} directly or through other objects:
        \begin{ffcode}
            goto
            [o1]
            ...
            |$\leadsto$|o1.forward o2
            ...
        \end{ffcode}
    \end{eodefinition}
    \begin{eodefinition}
        \emph{Backwarding goto} denotes an object \ff{goto}, which encapsulate object \ff{backward} directly or through other objects:
        \begin{ffcode}
            goto
            [o1]
            ...
            |$\leadsto$|o1.backward
            ...
        \end{ffcode}
    \end{eodefinition}
    \emph{Designations:}
    \begin{itemize}
        \item "$>$" is used to name attributes of an object, where on the left side of ">" is attribute itself and its name on the right side,
        \item "$[]$" is used for the specification of parameters of abstract objects,
        \item "$@$" is used to denote the object being decorated,
        \item "$\mapsto$" is used to denote an object, where free attributes stay in the parentheses on the left side and pairs, which represent bound attributes, stay on the right side, in double-square brackets,
        \item "$\leadsto$" denotes an arbitrary \emph{nesting} of an object, i.e. it can be encapsulated by an arbitrary number of objects.
    \end{itemize}

    \section{Sample}
    Let's consider the following example of code in EO, which uses GOTO object:

    \begin{ffcode}
        memory 0 > x
        [] > func
        goto > @
        [g]
        seq > @
        x.write (x.plus 1)
        if.
        x.lt 10
        g.backward
        QQ.io.stdout "X >= 10!\n"
        g.forward x
    \end{ffcode}

    \debate[Yegor]{Here I suggest we briefly explain the syntax of EO, since the language is not widely known. The reader will not understand what this code means and won't be able to digest it. Moreover, I would even suggest to also show a similar code in C++ language, so that the reader understand it better.

    ----------------Done}

    In this code snippet declares an \emph{atom}-object \ff{goto} with attribute \ff{g};
    mutable object \ff{memory}, which allows to reinitialize the bound object \ff{x} via \ff{.write} attribute;
    object \ff{eq}, which checks the equality of the objects passed to it;
    object \ff{if}, implementing behavior of branching;
    object \ff{seq}, which contains bound attributes as objects, which will be sequentially dataized;
    objects \ff{backward} and \ff{forward}, which are implementing behaviour of jumps: first of them starts dataization of object \ff{g} again, while the second one dataizes an object, passed as an attribute, and terminates dataization of object \ff{g}.

    Similar code snippet on C++ will be represented as:
    \begin{ffcode}
        int x = 0;
        int func() {
            l1:
            x = x + 1;
            if (x < 10)
            goto l1;
            else
            cout << "X >= 10!\n";
            return x;
        }
    \end{ffcode}

    We provide four independent transformations that replace \ff{goto} objects with a semantically equivalent analogue. After transformations, this example will appear as:
    \begin{ffcode}
        memory 0 > fl_1
        memory -1 > fl_2
        memory 0 > x
        [] > func
        seq > @
        fl_1.write 0
        fl_2.write -1
        while. > rem!
        or.
        fl_1.eq 0
        fl_2.eq -1
        [i]
        seq > @
        fl_1.write 1
        fl_2.write 0
        x.write (x.plus 1)
        if.
        x.lt 10
        fl_1.write 0
        QQ.io.stdout "X >= 10!\n"
        if.
        fl_1.eq 1
        fl_2.write 1
        TRUE
        if.
        fl_2.eq 1
        x
        rem
    \end{ffcode}

    Here, \ff{while} is an object, implementing behavior of loops: its first bound attribute implements condition of loop dataization and second one implies "body" of loop.

    On C++, the code snippet above after transformation will presented as:
    \begin{ffcode}
        int fl_1 = 0;
        int fl_2 = 0;
        int x = 0;
        int func() {
            fl_1 = 0;
            fl_2 = -1;
            do {
                fl_1 = 1;
                fl_2 = 0;
                x = x + 1;
                if (x < 10)
                fl_1 = 0;
                else
                cout << "X >= 10!\n";
                if (fl_1 == 1)
                fl_2 = 1;
            } while (fl_1 == 0 || fl_2 == -1);
            if (fl_2 == 1)
            return x;
        }
    \end{ffcode}

    Thus, the main purpose of our research is:
    \begin{enumerate}
        \item To formulate inference rules for the transformations that we propose,
        \item proof that these 4 transformations are necessary and sufficient to solve the problem under consideration,
        \item suggest a program tool, which
        enables automatic elimination of jump objects in programs written in EO.
    \end{enumerate}

@@

    Some objects in EO may need to be platform specific and can’t be composed from other existing objects — they are called \emph{atoms}.
    All jump statements are implemented as a
    single \emph{atom}-object \ff{goto}.
    The implementation of \ff{goto} in EO implies two types of jumps - forward jump and backward jump.
    It is required to cover all possible cases for both of them.

%%The proposed algorithm is a series of transformations that are performed depending on the type of "jump".
    In total, we describe 4 types of transformations:
    \begin{inparaenum}[1)]
    \item \emph{Simple goto} (further \emph{SG}),
    \item \emph{Goto forward/backward} (further \emph{GF}/\emph{GB}) and
    \item \emph{Terminating the while object} (further \emph{TW}).

    Transformations are independent and not sequential, and each one is executed in its own particular case.
    For example, \emph{SG} is only a "preparation" for the actual replacement of the jump by a semantically equivalent analog and is executed if there is no condition before the jump itself.
    \emph{GF}/\emph{GB} are the main transformations, as they contain the jump replacement itself and cover most possible cases of using forward and backward jumps.
    \emph{TW} is an "addition" for \emph{GF} and \emph{GB} transformations, since the behavior of the \ff{while} object in EO is different from all other objects and dataization of jump inside the loop requires separate consideration.
    \end{inparaenum}

%%Thus, the algorithm itself consists of the following steps:
%%\begin{inparaenum}[1)]
%%\item Apply \emph{SF}/\emph{SB} transformation if necessary,
%%\item Find all occurrences of forward/backward jump's calls,
%%\item Apply \emph{GF}/\emph{GB} transformations to all of them,
%%\item If there is a \emph{while.} object that wraps the jump call, apply the \emph{WF}/\emph{WB} transformation.
%%\end{inparaenum}

    The combination of all 4 transformations implies a complete algorithm for getting rid of the GOTO object, as they cover all possible cases of using GOTO and forward/backward jumps in the language.

    Since we consider a solution to the problem in strictly object-oriented programming languages, which means that the whole program is a set of objects where some objects can encapsulate others.
    It should be noted that each object in the program has its own level of \emph{nesting} relative to other objects.

    \debate[Yegor]{If you give a definition, make sure it's followed by example. Now it's not really clear what it means.

    -----------------Done}

    We will also consider our transformations in the form of $\varphi$-calculus, on which EO is based.
    The proposed $\varphi$-calculus represents an object model through data and objects, while operations with them are possible through abstraction, application and decoration.
    The calculus introduces a formal apparatus for manipulations with objects.

    \subsection{Simple goto} \label{SG}
    As mentioned above, this transformation is a "preparation" for \emph{GF}/\emph{GB} which are make the replacement of object \ff{goto} itself.
    Since \emph{GF} and \emph{GB} only consider the use of the \ff{forward}/\ff{backward} objects when they are directly encapsulated in the \ff{if} object(\cref{GF}/\cref{GB}), it is required to add an additional \ff{if} when it is absent before the jump.
    Thus, the transformation under consideration adds an additional fictitious condition $\emph{if (true)}$ if it necessary.

    Let's consider the case where the \ff{forward} is not directly encapsulated in the \ff{if} object:
    \begin{ffcode}
        goto
        [g]
        |$\leadsto$|x|$\label{ln:sf0}$|
        |$\alpha_1$|
        ...
        |$\alpha_n$|
        g.forward y
        |$\alpha_{n+1}$|
        ...
        |$\alpha_{m}$||$\label{ln:sf1}$|
    \end{ffcode}
    Here, $\alpha_{i}$, $x$ and $y$ defines any object in the language, excluding \ff{if}.

    Applying the transformation, the example above maps to:
    \begin{ffcode}
        goto
        [g]
        |$\leadsto$|x|$\label{ln:tsf0}$|
        |$\alpha_1$|
        ...
        |$\alpha_n$|
        if.
        TRUE
        g.forward y
        TRUE
        |$\alpha_{n+1}$|
        ...
        |$\alpha_{m}$||$\label{ln:tsf1}$|
    \end{ffcode}

    Describing transformation in terms of $\varphi$-calculus, code at lines ~\ref{ln:sf0}---\ref{ln:sf1} will be represented as:
    \begin{equation}
        \begin{split}
            & x(\alpha_1,\,\ldots\,,\,\alpha_n, \br
            & \quad \emph{g}.\ff{forward}( y ), \br
            & \quad \alpha_{n+1},\,\ldots\,,\,\alpha_m) \\
        \end{split}
    \end{equation}
    and code at lines ~\ref{ln:tsf0}---\ref{ln:tsf1} as:
    \begin{equation}
        \begin{split}
            & x(\alpha_1,\,\ldots\,,\,\alpha_n, \br
            & \quad \ff{TRUE}.\ff{if}(\emph{g}.\ff{forward}( y )), \br
            & \quad \alpha_{n+1},\,\ldots\,,\,\alpha_m) \\
        \end{split}
    \end{equation}

    \begin{eotheorem}
        Any object is semantically equivalent to the same object encapsulated in \ff{TRUE}.\ff{if}:
        \begin{equation*}
            \forall x (x \equiv \ff{TRUE}.\ff{if}(x)).
        \end{equation*}
    \end{eotheorem}

    \begin{proof}
        By definition of the language, if there is any object \emph{Z}, then the following equivalence follows:
        \begin{equation}
            Z \equiv if(true,\,Z,\,true),
        \end{equation}
        where first argument in \emph{if} is condition itself, second argument dataizes when condition is \emph{true} and third argument dataizes when condition is \emph{false}.
    \end{proof}

    \begin{equation*}
        \dfrac
        {\begin{split}
             & x(\alpha_1,\,\ldots\,,\,\alpha_n, \br
             & \quad \emph{g}.\ff{forward}( y ), \br
             & \quad \alpha_{n+1},\,\ldots\,,\,\alpha_m) \\
        \end{split}}
        {\begin{split}
             & x(\alpha_1,\,\ldots\,,\,\alpha_n, \br
             & \quad \ff{TRUE}.\ff{if}(\emph{g}.\ff{forward}( y )), \br
             & \quad \alpha_{n+1},\,\ldots\,,\,\alpha_m) \\
        \end{split}}
        \jrule{1}
    \end{equation*}

%%Note that in this example, \ff{forward} is simply an attribute of the abstract object \emph{g} and is dataized as soon as the queue reaches it.
    In such cases where there is an "unconditional" jump, first of all we transform it into "conditional" jump by encapsulating it in additional object \ff{if}---\emph{if (true)}.
    This will not violate the program logic, since added condition will always be \emph{TRUE} and "THEN" brunch of if-statement with object \ff{forward} will always be dataized.

    The same logic applies with the \ff{backward} object.

    \subsection{Goto forward} \label{GF}
    In this section, we discuss the transformation by replacing \emph{forwarding goto} with a semantic equivalent.
    Based on the concept of transformation, only cases in which \ff{forward} is directly encapsulated in the object \ff{if} are considered.

    The essence of the transformation is that we create an additional $flag$ object which will assign as "0" or "1" and will show whether the jump condition has been met.
    Instead of dataization \ff{forward}, we will dataize the object passed to it as an argument and reinitialize the $flag$ object.
    Moreover, it is necessary to protect from dataization all objects following the \ff{if} object containing the jump, and with \emph{nesting} not less than that of the goto object itself, since they should not be dataized when the jump condition is met.
    This requires encapsulating each such object with an additional object \ff{if} to check the equality $flag$ to an object \ff{TRUE} - i.e. whether the jump condition has been met before or not.

    Let's consider the following example:
    \begin{ffcode}
        goto
        [g]
        |$\leadsto$|x|$\label{ln:of0}$|
        |$\alpha_1$|
        ...
        |$\alpha_n$|
        if.
        k
        g.forward y
        |$\alpha_{n+1}$|
        |$\alpha_{n+2}$|
        ...
        |$\alpha_{m}$|
        |$\alpha_{m+1}$||$\label{ln:of1}$|
        ...
    \end{ffcode}
    Here, $\alpha_i$, $x$ and $y$ defines any object in the language, $k$ defines an object, which when dataized equals to a data object \ff{TRUE}.

    \debate[Yegor]{We don't have 'values' in EO. We only have objects. Moreover, we don't 'return' anything. The result of dataization may be equal to a data object 'TRUE'.}

    Applying the transformation, the example above maps
    to:
    \begin{ffcode}
        memory 0 > flag
        goto
        [g]
        |$\leadsto$|x|$\label{ln:tof0}$|
        |$\alpha_1$|
        ...
        |$\alpha_n$|
        if.
        k
        seq
        y
        flag.write 1
        |$\alpha_{n+1}$|
        if.
        eq. (flag 0)
        |$\alpha_{n+2}$|
        TRUE
        ...
        if.
        eq. (flag 0)
        |$\alpha_{m}$|
        TRUE
        if.
        eq. (flag 0)
        |$\alpha_{m+1}$|
        TRUE|$\label{ln:tof1}$|
        ...
    \end{ffcode}

    In terms of $\varphi$-calculus lines ~\ref{ln:of0}--\ref{ln:of1} appears as:
    \begin{equation}
        \begin{split}
            & x (\alpha_1,\,\ldots\,,\,\alpha_n, \br
            & \quad k.\ff{if}(\,\emph{g}.\ff{forward}( y ),\,\alpha_{n+1}),\br
            & \quad \alpha_{n+2},\,\ldots\,,\,\alpha_m), \\
            & \alpha_{m+1} \\
        \end{split}
    \end{equation}
    whereas the lines ~\ref{ln:tof0}--\ref{ln:tof1} will be represented as:
    \begin{equation}
        \begin{split}
            & x (\alpha_1,\,\ldots\,,\,\alpha_n, \br
            & \quad k.\ff{if}(\ff{seq}(y,\,flag.\ff{write}(1)),\,\alpha_{n+1}),\br
            & \quad flag.\ff{eq}(0).\ff{if}(\alpha_{n+2}),\,\ldots\,,\br
            & \quad flag.\ff{eq}(0).\ff{if}(\alpha_m)), \\
            & flag.\ff{eq}(0).\ff{if}(\alpha_{m+1}) \\
        \end{split}
    \end{equation}

    \begin{eotheorem}
        \emph{Forwarding goto} maps to semantically equivalent analog without \ff{forward} objects:
        \begin{equation*}
            \begin{split}
                & \ff{goto}(\llbracket g\,\mapsto\,\varnothing,\,\varphi\,\leadsto\,x(\ldots,\,\leadsto g.\ff{forward}(y),\,\ldots)\rrbracket) \\
                & \downarrow \\
                & \ff{goto}(\llbracket g\,\mapsto\,\varnothing,\,\varphi\,\leadsto\,x(\ldots,\,\leadsto y,\,\ldots)\rrbracket) \\
            \end{split}
        \end{equation*}
    \end{eotheorem}
